---
title: "EDAProblemSet3"
output: pdf_document
---
Load the required libraries

```{r}

require(gridExtra)
require(dplyr)
require(tidyr)
require(ggplot2)
```

Looking at the structure of the diamonds dataset

```{r}


require(datasets)
str(diamonds)

#unique(diamonds$color)
```

Plot the histogram for Price

```{r}
qplot(diamonds$price, 
      geom="histogram", 
     # binwidth=0.1, 
      main="Histogram for Price", 
      xlab="Price",
      fill=I("blue"), 
      col=I("red"), 
      alpha=I(.2),
    breaks=seq(500, 750,by=20),
      xlim = c(500,750))
```

Plot a graph for each type of cut and then arrange them in a grid.

```{r}
p1 <- qplot(diamonds[diamonds$cut == 'Fair',]$price, 
            geom="histogram", 
            # binwidth=0.1, 
            main="Histogram for Price", 
            xlab="Price",
            fill=I("blue"), 
            col=I("black"), 
            alpha=I(.2))

p2 <- qplot(diamonds[diamonds$cut == 'Good',]$price, 
            geom="histogram", 
            # binwidth=0.1, 
            main="Histogram for Price", 
            xlab="Price",
            fill=I("blue"), 
            col=I("black"), 
            alpha=I(.2))

p3 <- qplot(diamonds[diamonds$cut == 'Very Good',]$price, 
            geom="histogram", 
            # binwidth=0.1, 
            main="Histogram for Price", 
            xlab="Price",
            fill=I("blue"), 
            col=I("black"), 
            alpha=I(.2))

p4 <- qplot(diamonds[diamonds$cut == 'Premium',]$price, 
            geom="histogram", 
            # binwidth=0.1, 
            main="Histogram for Price", 
            xlab="Price",
            fill=I("blue"), 
            col=I("black"), 
            alpha=I(.2))

p5 <- qplot(diamonds[diamonds$cut == 'Ideal',]$price, 
            geom="histogram", 
            # binwidth=0.1, 
            main="Histogram for Price", 
            xlab="Price",
            fill=I("blue"), 
            col=I("black"), 
            alpha=I(.2))

grid.arrange(p1, p2, p3,p4,p5,ncol=2)

```

Implement the same graph, using facet_wrap functionality in ggplot and not have a fixed y-axis scale.

```{r}
qplot(x = price, data = diamonds) + facet_wrap(~cut, scales="free_y")
```

Plot a graph (boxplot) for the price/carat for each type of cut and save the graphs to a local directory.

```{r}
ggplot(diamonds, aes(cut,price/carat)) + geom_boxplot()
ggsave("box_by_cut.png")
ggplot(diamonds, aes(clarity,price/carat)) + geom_boxplot()
ggsave("box_by_clarity.png")
ggplot(diamonds, aes(color,price/carat)) + geom_boxplot()
ggsave("box_by_color.png")

```

Looking at the summary of Price for best and worst colors.

```{r}

summary(subset(diamonds, color=="D")$price)

summary(subset(diamonds, color=="J")$price)

IQR(subset(diamonds, color=="D")$price)

IQR(subset(diamonds, color=="J")$price)
```

Draw a frequency polygib to check the frequency of different weights of the diamonds.

```{r}
qplot(carat, data = diamonds, geom = "freqpoly", bins=100, xlim=c(2.0,5.0))

qplot(carat, data = diamonds, geom = "freqpoly", bins=100, xlim=c(2.0,3.0))

qplot(carat, data = diamonds, geom = "freqpoly", bins=100, xlim=c(1.8,2.0))

qplot(carat, data = diamonds, geom = "freqpoly", bins=100, xlim=c(1.5,1.6))

qplot(carat, data = diamonds, geom = "freqpoly", bins=100, xlim=c(1.0,1.1))

qplot(carat, data = diamonds, geom = "freqpoly", bins=100, xlim=c(0.5,0.8))

qplot(carat, data = diamonds, geom = "freqpoly", bins=100, xlim=c(0.1,0.5))

```


Below code is to answer the last set of questions in Problem Set 3.

Save the Unemployment data set for 15+ year olds for several countries for last 25+ years. The data is avaiable in Excel which was converted into CSV for the ease of reading into R.

```{r}

unm <- read.csv("unemployment_15.csv")

head(unm)

names(unm) <- c("country",1981:2005,"dummy")
```

reshape the data, convert the data from columns to rows. Filter the NAs after the conversion.

```{r}
unmg <- gather(unm, 'year',"unm",2:27)

unmg <- filter(unmg, !is.na(unm))



```

Plot the histrigrams for unemployment rates

```{r}
qplot(unmg$unm, geom="histogram", xlab = "unemployment", bins=100)

qplot(x=unm, data=unmg) + facet_wrap(~country, scales="free_y")

ggplot() + geom_line(data=unmg, aes(x=year,y=unm,group=country, color=country))
```

Check the plots for few countries

```{r}

ggplot() + geom_line(data=filter(unmg, country=="Spain"), aes(x=year,y=unm,group=country, color=country))


ggplot() + geom_line(data=filter(unmg, country=="Portugal"), aes(x=year,y=unm,group=country, color=country))

```
