---
title: "UDA651ProblemSet5"
author: "Mallesh"
date: "December 28, 2015"
output: pdf_document
---

Load the diamonds dataset and required libraries.

```{r}
require(dplyr)
require(ggplot2)
require(gridExtra)

data("diamonds")

```

Create a histogram of diamond prices, Facet the histogram by diamond color and 
use cut to color the histogram bars.

```{r}
qplot(data=diamonds, price, color=cut, bins=10) + facet_wrap(~color) +
  scale_fill_brewer(type = 'qual')
```

Create a scatterplot of diamond price vs. table and color the points by the cut of the diamond.

```{r}
ggplot(data=diamonds, aes(x=table, y=price)) + geom_point(aes(color=cut))

```

What is the typical table range for cut IDEAL. 

```{r}

range(filter(diamonds, cut=="Ideal")$table)

```

Create a scatterplot of diamond price vs.volume (x * y * z) and color the points by the clarity of diamonds. Use scale on the y-axis to take the log10 of price. You should also omit the top 1% of diamond volumes from the plot.

```{r}
ggplot(data=filter(diamonds, x*y*z >0), aes(x=x * y * z, y=price)) + 
  xlim(0, quantile(diamonds$x * diamonds$y * diamonds$z, 0.99)) +
  geom_point(aes(color=clarity)) +
  scale_y_log10()
```

create a new variable called 'prop_initiated' in the Pseudo-Facebook data set. The variable should contain the proportion of friendships that the user initiated.

```{r}

pf <- read.csv("pseudo_facebook.tsv", sep="\t")

pf <- pf %>%
  mutate(prop_initiated = ifelse(friend_count > 0 ,friendships_initiated/friend_count,NA))

#pf$prop_initiated <- ifelse(is.finite(pf$prop_initiated),pf$prop_initiated,0)
```

Create a line graph of the median proportion of friendships initiated ('prop_initiated') vs. tenure and color the line segment by year_joined.bucket.

```{r}

pf <- pf %>% mutate(year_joined = floor(2014 - (tenure/365) ))

pf$year_joined.bucket <- cut(pf$year_joined, 
                             breaks=c(2004, 2009,2011, 2012, 2014))

ggplot(data=pf, aes(x=tenure, y=prop_initiated)) +
  geom_line(aes(color=year_joined.bucket), stat="summary", fun.y=median) 

```
Smooth the last plot you created of of prop_initiated vs tenure colored by
year_joined.bucket. You can bin together ranges of tenure or add a smoother to the plot.

```{r}
ggplot(data=pf, aes(x=tenure, y=prop_initiated)) +
  geom_line(aes(color=year_joined.bucket), stat="summary", fun.y=median) +
  #geom_line(aes(color=year_joined.bucket), stat="summary", fun.y=mean)  
  geom_smooth()
```

```{r}
ggplot(data=pf, aes(x=tenure, y=prop_initiated)) +
  #geom_line(aes(color=year_joined.bucket), stat="summary", fun.y=median) +
  geom_line(aes(color=year_joined.bucket), stat="summary", fun.y=mean)  +
  geom_smooth()
```

Create a scatter plot of the price/carat ratio of diamonds. The variable x should be assigned to cut. The points should be colored
by diamond color, and the plot should be faceted by clarity.

```{r}
ggplot(data=diamonds, aes(x=cut, y=price/carat)) +
  geom_point(aes(color=color)) + facet_wrap(~clarity)
```

Below is the final exercise where a dataset from Gapminder is downloaded.

Save the Unemployment data set for 15+ year olds for several countries for last 25+ years. The data is avaiable in Excel which was converted into CSV for the ease of reading into R.

```{r}
require(tidyr)

unm <- read.csv("unemployment_15.csv")

head(unm)

names(unm) <- c("country",1981:2005,"dummy")
```

reshape the data, convert the data from columns to rows. Filter the NAs after the conversion.

```{r}
unmg <- gather(unm, 'year',"unm",2:27)

unmg <- filter(unmg, !is.na(unm))

```

create bins for each decade:

```{r}
unmg <- unmg %>%
  mutate(year_num = as.integer(year) + 1980)


unmg$year.bucket <- cut(unmg$year_num, 
                             breaks=c(1980,1991,2000,2005), right=T)
```

plot unemplyment mean across the years:

```{r}
ggplot(unmg, aes(x=year, y=unm)) + geom_point(aes(color=year.bucket))

```

Scatter plot doesn't really show any patterns except that the minimum employment rate increased across the decades.

Let us confirm our observation by pointing the median unemplyment rate.

```{r}
ggplot(unmg, aes(x=year_num, y=unm)) + geom_point(aes(color=year.bucket)) +
  
  geom_line(stat="summary", fun.y=median)
```
lets add facet wrap for different countries:

```{r}
ggplot(unmg, aes(x=year_num, y=unm)) + geom_point(aes(color=year.bucket)) +
  
  geom_line(stat="summary", fun.y=median) + facet_wrap(~country)
```

This graph indicates trend across countries. Advanced countries like Australia, Canada, United states show a downward trend, while some of the countries like China indicates an upward trend in unemplyment rate.