---
title: "UD651-Lesson3"
output: pdf_document
---

This document includes all the assignments done as part of the UD651 Lesson 3 course on Udacity.

Dataset used is a Pseudo facebook data downloaded from the course site and saved in the current working directory. 
https://s3.amazonaws.com/udacity-hosted-downloads/ud651/lesson3_student.rmd

Loading required libraries:

```{r}
require(ggplot2)
require(gridExtra)

```

```{r}
fb <- read.csv("pseudo_facebook.tsv", sep="\t")
```

Creating a histogram for the friend count separated by Gender:

```{r}
qplot(x = friend_count, data = na.omit(fb), binwidth = 25) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

Generating the summary data:

```{r}
table(fb$gender)

by(fb$friend_count, fb$gender, summary)
```


Analyze the tenure and its spread in the dataset. Converting the data into years since its in days:

```{r}

qplot(x = tenure/365, data = na.omit(fb), binwidth = 1)
```

Explore several scaling functions in ggplot and comparing them to the original plot.

```{r}
p1 <- qplot(x = friend_count, data = na.omit(fb), binwidth = 5) + 
  scale_x_continuous(limits = c(0, 150), breaks = seq(0, 150, 5))

p2 <- qplot(x = log10(friend_count+1), data = na.omit(fb), binwidth = 1) 

p3 <- qplot(x = friend_count, data = na.omit(fb), binwidth = 1) 

```

Arrange the above graphs in a grid:

```{r}

grid.arrange(p1, p2, p3, ncol=1)
```

Plotting a frequency polygon using the geom options on different scales

```{r}
qplot(x = www_likes, data = na.omit(fb), geom="freqpoly", color=gender) +
  scale_x_continuous(limits = c(0, 150), breaks = seq(0, 150, 5)) 

qplot(x = www_likes, data = na.omit(fb), geom="freqpoly", color=gender) +
  scale_x_log10()
```

Using the ylim operator to limit the Y limits

```{r}

qplot(x=gender, y = friend_count, data = subset(fb, !is.na(gender)), geom="boxplot",
      ylim = c(0,1000))
```

Using coord cartesian to limit the y values on friendships_initiated:

```{r}

qplot(x=gender, y = friendships_initiated, data = subset(fb, !is.na(gender)), geom="boxplot") +
      coord_cartesian(ylim = c(0,130))
```

Calculate the mobile_check_in s:

```{r}
fb$mobile_check_in <- NA

fb$mobile_check_in <- ifelse(fb$mobile_likes >0,1,0)

summary(fb$mobile_check_in)


sum(fb$mobile_check_in == 1)/
length(fb$userid)

```